Elimination Of Jump Objects

### 2022-06-30 paper/skeleton.tex
Update skeleton.tex

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false]{biblatex}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xcolor}

@@

\newtheorem{definition}{Definition}[section]

@@

@@

The solution we propose contains two types of transformations - forward jump and backward jump.
The algorithm itself consists in applying these transformations to replace both types of jumps.

\vspace{5mm}

@@

\item $\mu$ - Denotes an object or a chain of nested objects that return a boolean value (TRUE/FALSE) when executed.
\end{itemize}

@@

We will also look at our transformations in the form of $\varphi$-calculus, on which EO is based. The proposed $\varphi$-calculus represents an object model through data and objects, while operations with them are possible through abstraction, application, and decoration.
The calculus introduces a formal apparatus for manipulations with objects.

$\varphi$-calculus uses an alternative “arrow notation” to denote an object in a more compact way, where free attributes of an abstract object stay in the parentheses on the left side of the mapping symbol $\mapsto$ and pairs, which represent bound attributes, stay on the right side, in double-square brackets.

@@

\subsubsection{Simple Forward}
Consider the simplest example of using goto-forward:

\begin{ffcode}
goto
  [g]
    g.forward |\dlt{1}| > @
\end{ffcode}

In terms of $\varphi$-calculus, this code will appear as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \varphi \mapsto \ff{g}.\ff{forward} (\delta_1) \rrbracket ) \\
\end{split}
\end{equation}

@@

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing,\br
& \quad \varphi \mapsto \ff{if}.( TRUE, \ff{g}.\ff{forward}(\delta_1), TRUE ) \rrbracket)\\
\end{split}
\end{equation}

Same example on EO:

@@

\subsubsection{Goto forward wrapped in an object}

After the conversion described in section 3.1.1, the simplest call of forward jump will be presented as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \br
& \quad \varphi \mapsto \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \br
& \quad \varphi \mapsto \ff{if}.(\mu_1.\ff{not}, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \br
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \br
& \quad \varphi \mapsto \ff{if}.(\mu_1.\ff{not}, \delta_2, \delta_1) \rrbracket) \br
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{if.}(\mu_1.\ff{not}, \delta_2, \delta_1) \br
\end{split}
\end{equation}

\begin{theorem}
Expression 3 is semantically equivalent to expression 6:
\end{theorem}

\begin{equation*}
\dfrac
    {\begin{matrix}
    \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
    \varphi \mapsto \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket)
    \end{matrix}}
    {\ff{if.}(\mu_1.\ff{not}, \delta_2, \delta_1)}
    \jrule{1}
\end{equation*}

@@

Now let's look at an example program with nested objects containing goto-forward:

@@

\subsubsection{Goto backward wrapped in an object}
After the conversion described in Section 3.2.1, the simplest call of backward jump will be presented as:

>>>>

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false]{biblatex}
\usepackage{debate}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xcolor}

@@

\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[theorem]

@@

The proposed algorithm is a series of transformations that are performed depending on the type of "jump".
In total, we describe 6 types of transformations: \emph{Simple forward}/\emph{backward} (further \emph{SF}/\emph{SB}), \emph{Goto forward/backward wrapped by object} (further \emph{OF}/\emph{OB}) and \emph{Goto forward/ backward inside the while-loop} (further \emph{WF}/\emph{WB}).
Transformations are not sequential and independent, and each one is executed in its own particular case.
For example, \emph{SF}/\emph{SB} is only a "preparation" for the actual replacement of the jump by a semantically equivalent analog and is executed if there is no condition before the jump itself.
\emph{OF}/\emph{OB} is the main transformation, as it contains the jump replacement itself and covers most possible cases of calling forward or backward jump.
\emph{WF}/\emph{WB} is a "addition" for the previous transformation, since the behavior of the \emph{while.} object in EO is different from all other objects and the call of the jump inside the loop requires separate consideration.

Thus, the algorithm itself consists of the following steps: 
\begin{inparaenum}[1)]
\item Apply \emph{SF}/\emph{SB} transformation if necessary,
\item Find all occurrences of forward/backward jump's calls,
\item Apply \emph{OF}/\emph{OB} transformations to all of them,
\item If there is a \emph{while.} object that wraps the jump call, apply the \emph{WF}/\emph{WB} transformation.
\end{inparaenum}

The combination of all 6 transformations makes a complete algorithm for getting rid of the GOTO object, as they cover all possible cases of using GOTO and forward/backward jumps in the language.

@@

@@

\item $\mu$ - Denotes a not-empty object or a chain of nested objects that return a boolean value (TRUE/FALSE) when accessed.
\end{itemize}

@@

We will also consider our transformations in the form of $\varphi$-calculus, on which EO is based. The proposed $\varphi$-calculus represents an object model through data and objects, while operations with them are possible through abstraction, application, and decoration.
The calculus introduces a formal apparatus for manipulations with objects.

%%$\varphi$-calculus uses an alternative “arrow notation” to denote an object in a more compact way, where free attributes of an abstract object stay in the parentheses on the left side of the mapping symbol $\mapsto$ and pairs, which represent bound attributes, stay on the right side, in double-square brackets.

@@

\subsubsection{Simple Forward}

The \ff{goto} object may be used for unconditional forward jumps. In general case, covering most of the cases of calling goto-forward, it appears as:

\begin{ffcode}
goto
  [g]
    |\sgm{1}| > @
      |\sgm{2}|
      g.forward |\dlt{1}|
      |\sgm{3}|
\end{ffcode}

Here, \dlt{1} is any object, which will become the result of dataization of the \ff{goto} object, $\sigma_1$, $\sigma_2$ are ob
In terms of $\varphi$-calculus, this code may be expressed as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \br
& \quad \quad \quad \varphi \mapsto \sigma_1( \sigma_2, \ff{g}.\ff{forward} (\delta_1), \sigma_3) \rrbracket ) \\
\end{split}
\end{equation}

Where in parentheses are denoted free attributes, in square brackets are denoted bound attributes and the symbol $\mapsto$ denotes application/decoration of an object.

@@

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \varphi \mapsto \sigma_1 (\sigma_2, \br
& \quad \quad \quad \ff{if}.( TRUE, \ff{g}.\ff{forward}(\delta_1), TRUE ), \sigma_3) \rrbracket)\\
\end{split}
\end{equation}

Same example in EO:

@@

\subsubsection{Goto forward wrapped by object}

After the transformation described in section 3.1.1, the general case of calling forward jump will be presented as:

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \varphi \mapsto \sigma_1 (\sigma_2, \br
& \quad \quad \quad \ff{if}.( \mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2 ), \sigma_3) \rrbracket)\\
\end{split}
\end{equation}

\begin{theorem}
A goto object that contains an object or a chain of nested objects that wrap the \emph{.forward} attribute can be converted to a similar one, but without using the \emph{.forward} attribute.
\end{theorem}

In terms of $\varphi$-calculus, it can be represented as the following inference rule:

\begin{equation*}
\dfrac
    {\begin{matrix}
    & \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \varphi \mapsto \sigma_1 (\sigma_2, \\
    & \quad \quad \quad \quad \ff{if}.( \mu_1,     \ff{g}.\ff{forward}(\delta_1), \delta_2 ), \sigma_3) \rrbracket)\\
    \end{matrix}}
    {\begin{matrix}
    & \sigma_1(\sigma_2, \ff{if.}(\mu_1.\ff{not}, \delta_2, \delta_1), \\
    & \quad \quad \ff{if.}(\mu_1.\ff{not}, \sigma_3, TRUE)) \\
    \end{matrix}}
    \jrule{1}
\end{equation*}

\begin{proof}
To get rid of using ".forward" attribute, it is required to emulate work of jump itself.

First of all, let's invert condition of jump, swap objects in "THEN"/"ELSE" brunches of if-statement and remove \emph{.forward} attribute:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \br
& \quad \quad \varphi \mapsto \sigma_1 ( \sigma_2,  \ff{if}.(\mu_1.not, \delta_2, \delta_1), \sigma_3) \rrbracket) \\
\end{split}
\end{equation}

Since the jump does not execute all statements following the jump itself, we need to protect their execution by checking the jump condition for each of them:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \br
& \quad \quad \varphi \mapsto \sigma_1(\sigma_2,  \ff{if}.(\mu_1.\ff{not}, \delta_2, \delta_1), \br
& \quad \quad \quad \ff{if.}(\mu_1.\ff{not}, \sigma_3, TRUE)) \rrbracket) \\
\end{split}
\end{equation}

Thus, all objects that would not have executed in the presence of the \emph{.forward} attribute will not be executed now. And objects that were returned in branches of the old if-statement will also be executed under the same condition.

\end{proof}

\begin{corollary}
If there are no \emph{.forward} attributes inside the abstract object goto, no jumps will be performed and therefore it is possible to delete the abstract object goto itself without losing logic.
\end{corollary}

\begin{equation}
\begin{split}
& \sigma_1( \sigma_2, \ff{if}.(\mu_1.\ff{not}, \delta_2, \delta_1), \br
& \quad \quad \ff{if.}(\mu_1.\ff{not}, \sigma_3, TRUE)) \\
\end{split}
\end{equation}

Let's see, how this transformation will be presented in EO.

@@

Now let's look at an example of transformation, if the code fragment \emph{X} is wrapped into an object:

@@

\subsubsection{Goto backward wrapped by object}
After the conversion described in Section 3.2.1, the simplest call of backward jump will be presented as:
